\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-12-14 adaptive-estimation.tex
Starting to work on distributed part

\begin{figure}
  %%%%% COMMENTED TIKZ PICTURE %%%%%
\begin{center}
%% \begin{tikzpicture}[scale = 0.6]
%% \begin{axis}[
%% width=10cm, height=6cm,
%% xmin = 0, xmax=800, 
%% restrict y to domain = 0:3,
%% ymin = 0,
%% ymax = 3.4,
%% samples=10, 
%% xlabel= $n$,
%% ylabel = {$n\mathbb E \left[\left(\theta - {\theta}_n \right)^2 \right]$},
%% %xtick={-3,-2,-1,0,1,2,3},
%% %xticklabels={-3,-2,-1,0,1,2,3},
%% ytick={0,1,1.57},
%% yticklabels={0,1,$\pi/2$},
%% line width=1.0pt,
%% mark size=1.5pt,
%% ymajorgrids,
%% xmajorgrids,
%% legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
%% ]

%% \addplot[color = blue, solid, smooth] plot table [x = itr, y = SGD, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};
%% \addlegendentry{asymptotically optimal};

%% %\addplot[color = red, solid, smooth] plot table [x = itr, y = Bayes, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};\addlegendentry{one step optimal};

%% \addplot[color = red, solid, smooth] plot table [x = itr, y = split, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};
%% \addlegendentry{single interaction};

%% \end{axis}
%% \end{tikzpicture}
\caption{Normalized empirical risk versus number of samples $n$ for $10,000$ Monte Carlo trials with $f(x)$ the standard normal density. In each trial, $\theta$ is chosen uniformly over the interval $(-1.64,1.64)$. The single interaction strategy uses $n_1 = \lfloor \sqrt{n} \rfloor$ samples for its first stage. 
\label{fig:adaptive_error}  }
\end{center}
\end{figure}

>>>>

\begin{figure}
  %%%%% COMMENTED TIKZ PICTURE %%%%%
\begin{center}
\begin{tikzpicture}[scale = 0.6]
\begin{axis}[
width=10cm, height=6cm,
xmin = 0, xmax=800, 
restrict y to domain = 0:3,
ymin = 0,
ymax = 3.4,
samples=10, 
xlabel= $n$,
ylabel = {$n\mathbb E \left[\left(\theta - {\theta}_n \right)^2 \right]$},
%xtick={-3,-2,-1,0,1,2,3},
%xticklabels={-3,-2,-1,0,1,2,3},
ytick={0,1,1.57},
yticklabels={0,1,$\pi/2$},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]

\addplot[color = blue, solid, smooth] plot table [x = itr, y = SGD, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};
\addlegendentry{asymptotically optimal};

%\addplot[color = red, solid, smooth] plot table [x = itr, y = Bayes, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};\addlegendentry{one step optimal};

\addplot[color = red, solid, smooth] plot table [x = itr, y = split, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};
\addlegendentry{single interaction};

\end{axis}
\end{tikzpicture}
\caption{Normalized empirical risk versus number of samples $n$ for $10,000$ Monte Carlo trials with $f(x)$ the standard normal density. In each trial, $\theta$ is chosen uniformly over the interval $(-1.64,1.64)$. The single interaction strategy uses $n_1 = \lfloor \sqrt{n} \rfloor$ samples for its first stage. 
\label{fig:adaptive_error}  }
\end{center}
\end{figure}

### 2019-12-14 distributed-estimation.tex
Starting to work on distributed part

\begin{figure}
  %%%%% COMMENTED TIKZ PICTURE %%%%%
  \begin{center}
\begin{tikzpicture}[scale = 1]
\begin{axis}[
width=8cm, height=6cm,
xmax=2.5, 
xmode = log,
restrict y to domain = 0:100,
ymin = 0,
ymax = 1,
samples=1, 
xlabel= {$\sigma/T$},
%xtick={-0.5,0,0.5},
%xticklabels={-$$, 0, $b$},
ytick={0,0.637,1},
yticklabels={0,$2/\pi$,1},
ylabel = {$\ARE$},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = red, smooth, line width = 1pt] plot table [x = x, y = y, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
\addlegendentry{optimal threshold density};

\addplot[color = black!35!green, smooth, dashed, line width = 1pt] plot table [x = x, y = z, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
\addlegendentry{uniform threshold desnity};
\end{axis}
\end{tikzpicture}
\caption{\label{fig:minimax_ARE} Minimax ARE versus $\sigma/T$ for $f(x) = \Ncal(\theta,\sigma^2)$  and $\theta \in \Theta = [-T,T]$. The dashed curve (green) is the ARE under a uniform threshold density over $\Theta$ given by $K_{\unif}\sigma^2$, where $\kappa_{\unif}$ is given by \eqref{eq:uniform_risk}. }.
\end{center}
\end{figure}

%We have seen that using threshold detection, an ARE $\eta(0)\sigma^2$ can only be attained at a single point in $\Theta$. Next, we prove that with any set of messages obtained via a distributed strategy, there exists at least one $\theta \in \Theta$ such that the estimator of $\theta$ has ARE strictly smaller than $\eta(0)\sigma^2$.

\begin{figure}
\begin{center}
%
\begin{tikzpicture}[scale = 0.55]
\begin{axis}[
width=8cm, height=6cm,
xmin = -0.5, xmax=0.5, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 1,
samples=10, 
xlabel= {$dt$},
title = {$\sigma/\sigma_{\theta} = 4$},
xtick={-0.5,0,0.5},
xticklabels={-$T$, 0, $T$},
ytick={0,1},
ylabel = {$\lambda(dt)$},
yticklabels={0,1},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig4.csv};
%\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
\end{axis}
\end{tikzpicture}
%
\begin{tikzpicture}[scale = 0.55]
\begin{axis}[
width=8cm, height=6cm,
xmin = -0.5, xmax=0.5, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 1,
samples=10, 
xlabel= {$dt$},
xtick={-0.5,0,0.5},
title = {$\sigma/\sigma_{\theta} = 3$},
xticklabels={-$T$, 0, $T$},
ytick={0,1},
%ylabel = {$\lambda(dt)$},
yticklabels={0,1},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig3.csv};
%\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
\end{axis}
\end{tikzpicture}
%
\begin{tikzpicture}[scale = 0.55]
\begin{axis}[
width=8cm, height=6cm,
xmin = -0.5, xmax=0.5, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 1,
samples=10, 
xlabel= {$dt$},
title = {$\sigma/\sigma_{\theta} = 2$},
xtick={-0.5,0,0.5},
xticklabels={-$T$, 0, $T$},
ytick={0,1},
ylabel = {$\lambda(dt)$},
yticklabels={0,1},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig2.csv};
%\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
\end{axis}
\end{tikzpicture}
%
\begin{tikzpicture}[scale = 0.55]
\begin{axis}[
width=8cm, height=6cm,
xmin = -0.5, xmax=0.5, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 1,
samples=10, 
xlabel= {$dt$},
xtick={-0.5,0,0.5},
xticklabels={-$T$, 0, $T$},
title = {$\sigma/\sigma_{\theta} = 1$},
ytick={0,1},
%ylabel = {$\lambda(dt)$},
yticklabels={0,1},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig1.csv};
%\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
\end{axis}
\end{tikzpicture}
\caption{\label{fig:opt_density}
Optimal threshold density $\lambda^\star(dt)$ that minimizes the asymptotic Bayes risk \eqref{eq:cvx_average} for a uniform prior with $\sigma/\sigma_\theta=1,2,3,4$, where $\sigma_\theta^2=T^2/3$ is the variance of the prior. 
}
\end{center}
\end{figure}

@@

\iffalse
\begin{figure}
\begin{center}
\begin{tikzpicture}[scale = 1]
\begin{axis}[
width=8cm, height=6cm,
xmin = 0, xmax=2, 
restrict y to domain = 0:100,
ymin = 0,
ymax = 10,
samples=1, 
xlabel= {$\sigma/\sigma_\theta$},
%xtick={-0.5,0,0.5},
%xticklabels={-$$, 0, $b$},
ytick={0,1.5708,5},
yticklabels={0,$\pi/2$,5},
ylabel = {$R_{\pi}^\star/\sigma^2$},
line width=1.0pt,
mark size=1.5pt,
ymajorgrids,
xmajorgrids,
legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
]
\addplot[color = red, smooth] plot table [x = x, y = Runif, col sep=comma] {./Figs/unif_Bayes_Risk.csv};
%\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
\addplot[color = red, smooth, dashed, line width = 0.5pt] plot table [x = x, y = Rbound, col sep=comma] {./Figs/unif_Bayes_Risk.csv};
\addplot[color = black, dashed] {3.14159 / 2};
\end{axis}
\end{tikzpicture}
\caption{Asymptotic Bayes risk $R_{\pi}^\star$ in estimating the mean of a normal distribution ($P_X = \Ncal(\theta, \sigma^2)$ under an optimal threshold distribution $\lambda^\star$, when the prior $\pi$ is the uniform distribution over $\Theta = [-0.5,0.5]$. 
%
%versus $\sigma/\sigma_\theta = 2\sqrt{3}\sigma$.
The distribution $\lambda^\star$ is the minimizer of \eqref{eq:cvx_average}. It is illustrated for various cases in Fig.~\ref{fig:opt_density}. The dashed curve represents the upper bound \eqref{eq:upper_bound}. 
\label{fig:dist_bound_uniform}  }
\end{center}
\end{figure}

As can be seen in Fig.~\ref{fig:opt_density} for the case $P_X = \Ncal(\theta ,\sigma^2)$ and a uniform $\pi$, when the radius of $\Theta$ is small compared to $\sigma$, the optimal distribution $\lambda^\star$ is a mass distribution. In this case, the ML estimator reduces to the estimator ${\theta}_n$ of \eqref{eq:estimator_naive}. As the following proposition shows, the Bayes risk for this choice of $\lambda$ is maximal, and thus provides an upper bound on the Bayes risk under any $\lambda$. 
 \eqref{eq:cvx_average}. 
\begin{prop}\label{prop:upper_bound}
For any prior $\pi(d\theta)$ and $\theta_0$ in the support of $\eta(x)$ we have
\begin{equation} 
\label{eq:upper_bound}
R_\pi^\star  \leq 
 \int \frac{\pi(d\theta)}{\eta \left( \theta_0 - \theta \right)},
\end{equation}
Furthermore, assuming that $\theta_0 = \ex{\theta}$ and that $\pi$ has a finite third moment $\sigma_\theta^3$, we have: 
\begin{equation}
\label{eq:bound_Taylor}
R^\star_\pi \leq \frac{1}{4 f^2(0)} + \left(\frac{1}{4 f^2(0)} \frac{-f''(0)}{f(0)} -1 \right) \sigma_\theta^2 + O(\sigma_\theta^3).
\end{equation}
\end{prop}
\begin{proof}
The function $x \rightarrow 1/x$ is convex for positive values, hence Jensen's inequality implies
\[
\left( \int \eta \left( t-\theta\right) \lambda(dt) \right)^{-1}  \leq \int  \frac{ \lambda(dt)}{ \eta \left( t-\theta\right)  }. 
\]
Therefore, the expected value of $\kappa^{-1}(\theta)$ satisfies
\begin{align}
\ex{  \frac{1}{\kappa(\theta)}}  \leq \int \int \frac{\pi(d\theta) \lambda(dt) }{\eta \left( t - \theta \right)}. \label{eq:upper_bound_proof}
\end{align}
The bound \eqref{eq:upper_bound} is obtained by taking $\lambda$ to be a mass distribution at any $\theta_0$ in the support of $\eta(x)$. Finally, \eqref{eq:bound_Taylor} is obtained by expanding $1/\eta(x)$ to a third order Taylor series around zero
%\[
%1/\eta(x) = \frac{1}{4f^2(0)}  + \left(\frac{1}{4 f^2(0)} \frac{-f''(0)}{f(0)} -1 \right)  x^2 + O(x^3),
%\]
and taking its expectation with respect to $\pi$ at $x=\theta_0-\theta$. 
\end{proof}
We note that the function $1/\eta(x)$ is quasi-convex and symmetric around zero, so the choice $\theta_0 = \ex{ \theta}$ minimizes the RHS of \eqref{eq:upper_bound} among all $\theta_0$ in the support of $\eta(x)$.\par
The bound \eqref{eq:upper_bound} is not trivial as long as the integral in the RHS of \eqref{eq:upper_bound} is finite, i.e., whenever the tail of $\pi(\theta)$ vanishes fast enough compared to $\eta(x)$. The expansion \eqref{eq:bound_Taylor} implies that this bound becomes tight whenever the support of the optimal distribution is a mass distribution at $\mathbb E \theta$, in which case the expected value of $\kappa^{-1}(\theta)$ approaches $1/\eta(0) = 1/4f^2(0)$.

\begin{example} \label{ex:bound}
In the normal case $P_X = \Ncal(\theta, \sigma^2)$, the bound in \eqref{eq:bound_Taylor} implies 
\[
\frac{R^\star_\pi}{\sigma^2}  \leq \frac{\pi}{2} + \left( \frac{\pi}{2} -1 \right) \left( \frac{ \sigma_\theta}{\sigma} \right)^2 + O \left(  \frac{\sigma_\theta} { \sigma} \right)^3. 
\]
It follows that the ARE approaches its maximal value of $2/\pi$ whenever $\sigma_\theta/\sigma$ is small. The exact value of \eqref{eq:upper_bound} in this case, as well as the Bayes ARE with the optimal threshold density $\lambda^\star$ for a uniform $\pi$, are illustrated in Fig.~\ref{fig:dist_bound_uniform}. 
%i.e., each encoder simply reports "$1$" or "$-1$" whenever $X_i$ is larger or smaller than $\theta_0$, respectively. Intuitively, an accurate estimation of $\theta$ in this case is possible only if a sufficient mix of ''$1$''s and ''$-1$''s is obtained from the sample. When $\sigma_\theta^2$ is high compared to $\simga^2$, one of the events $"1"$ or "$-1"$ becomes too rare to
\end{example}
\fi

>>>>

\begin{figure}
  %%%%% COMMENTED TIKZ PICTURE %%%%%
  \begin{center}
%% \begin{tikzpicture}[scale = 1]
%% \begin{axis}[
%% width=8cm, height=6cm,
%% xmax=2.5, 
%% xmode = log,
%% restrict y to domain = 0:100,
%% ymin = 0,
%% ymax = 1,
%% samples=1, 
%% xlabel= {$\sigma/T$},
%% %xtick={-0.5,0,0.5},
%% %xticklabels={-$$, 0, $b$},
%% ytick={0,0.637,1},
%% yticklabels={0,$2/\pi$,1},
%% ylabel = {$\ARE$},
%% line width=1.0pt,
%% mark size=1.5pt,
%% ymajorgrids,
%% xmajorgrids,
%% legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
%% ]
%% \addplot[color = red, smooth, line width = 1pt] plot table [x = x, y = y, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
%% \addlegendentry{optimal threshold density};

%% \addplot[color = black!35!green, smooth, dashed, line width = 1pt] plot table [x = x, y = z, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
%% \addlegendentry{uniform threshold desnity};
%% \end{axis}
%% \end{tikzpicture}
%% \caption{\label{fig:minimax_ARE} Minimax ARE versus $\sigma/T$ for $f(x) = \Ncal(\theta,\sigma^2)$  and $\theta \in \Theta = [-T,T]$. The dashed curve (green) is the ARE under a uniform threshold density over $\Theta$ given by $K_{\unif}\sigma^2$, where $\kappa_{\unif}$ is given by \eqref{eq:uniform_risk}. }.
%% \end{center}
%% \end{figure}

%% %We have seen that using threshold detection, an ARE $\eta(0)\sigma^2$ can only be attained at a single point in $\Theta$. Next, we prove that with any set of messages obtained via a distributed strategy, there exists at least one $\theta \in \Theta$ such that the estimator of $\theta$ has ARE strictly smaller than $\eta(0)\sigma^2$.

%% \begin{figure}
%% \begin{center}
%% %
%% \begin{tikzpicture}[scale = 0.55]
%% \begin{axis}[
%% width=8cm, height=6cm,
%% xmin = -0.5, xmax=0.5, 
%% restrict y to domain = 0:100,
%% ymin = 0,
%% ymax = 1,
%% samples=10, 
%% xlabel= {$dt$},
%% title = {$\sigma/\sigma_{\theta} = 4$},
%% xtick={-0.5,0,0.5},
%% xticklabels={-$T$, 0, $T$},
%% ytick={0,1},
%% ylabel = {$\lambda(dt)$},
%% yticklabels={0,1},
%% line width=1.0pt,
%% mark size=1.5pt,
%% ymajorgrids,
%% xmajorgrids,
%% legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
%% ]
%% \addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig4.csv};
%% %\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
%% \end{axis}
%% \end{tikzpicture}
%% %
%% \begin{tikzpicture}[scale = 0.55]
%% \begin{axis}[
%% width=8cm, height=6cm,
%% xmin = -0.5, xmax=0.5, 
%% restrict y to domain = 0:100,
%% ymin = 0,
%% ymax = 1,
%% samples=10, 
%% xlabel= {$dt$},
%% xtick={-0.5,0,0.5},
%% title = {$\sigma/\sigma_{\theta} = 3$},
%% xticklabels={-$T$, 0, $T$},
%% ytick={0,1},
%% %ylabel = {$\lambda(dt)$},
%% yticklabels={0,1},
%% line width=1.0pt,
%% mark size=1.5pt,
%% ymajorgrids,
%% xmajorgrids,
%% legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
%% ]
%% \addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig3.csv};
%% %\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
%% \end{axis}
%% \end{tikzpicture}
%% %
%% \begin{tikzpicture}[scale = 0.55]
%% \begin{axis}[
%% width=8cm, height=6cm,
%% xmin = -0.5, xmax=0.5, 
%% restrict y to domain = 0:100,
%% ymin = 0,
%% ymax = 1,
%% samples=10, 
%% xlabel= {$dt$},
%% title = {$\sigma/\sigma_{\theta} = 2$},
%% xtick={-0.5,0,0.5},
%% xticklabels={-$T$, 0, $T$},
%% ytick={0,1},
%% ylabel = {$\lambda(dt)$},
%% yticklabels={0,1},
%% line width=1.0pt,
%% mark size=1.5pt,
%% ymajorgrids,
%% xmajorgrids,
%% legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
%% ]
%% \addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig2.csv};
%% %\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
%% \end{axis}
%% \end{tikzpicture}
%% %
%% \begin{tikzpicture}[scale = 0.55]
%% \begin{axis}[
%% width=8cm, height=6cm,
%% xmin = -0.5, xmax=0.5, 
%% restrict y to domain = 0:100,
%% ymin = 0,
%% ymax = 1,
%% samples=10, 
%% xlabel= {$dt$},
%% xtick={-0.5,0,0.5},
%% xticklabels={-$T$, 0, $T$},
%% title = {$\sigma/\sigma_{\theta} = 1$},
%% ytick={0,1},
%% %ylabel = {$\lambda(dt)$},
%% yticklabels={0,1},
%% line width=1.0pt,
%% mark size=1.5pt,
%% ymajorgrids,
%% xmajorgrids,
%% legend style= {at={(1,1)},anchor=north east,draw=black,fill=white,align=left}
%% ]
%% \addplot[color = blue, smooth, mark = o ] plot table [col sep=comma] {./Figs/unif_Bayes_lmd_b0.5_sig1.csv};
%% %\addlegendentry{$\frac{\sigma^2}{\sigma_\theta^2} = 1$};
%% \end{axis}
%% \end{tikzpicture}
\caption{\label{fig:opt_density}
Optimal threshold density $\lambda^\star(dt)$ that minimizes the asymptotic Bayes risk \eqref{eq:cvx_average} for a uniform prior with $\sigma/\sigma_\theta=1,2,3,4$, where $\sigma_\theta^2=T^2/3$ is the variance of the prior. 
}
\end{center}
\end{figure}

@@
